A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal

### 2020-08-12 Manuscript/PNAS-Adams-01-EKB.tex
vsn 2

\begin{abstract}
Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, however, analytical tools for comparing the
strength of that signal across traits remain largely underdeveloped.
Here we evaluate the efficacy of of Pagel's \(\lambda\) to correctly
estimate the strength of phylogenetic signal in phenotypic traits across
a range of input values. We find that \(\lambda\) behaves as a Bernoulli
random variable, where estimates are increasingly skewed at larger and
smaller input levels of phylogenetic signal. Further, the precision of
\(\lambda\) varies with input signal. Another measure, Blomberg's
\(\kappa\), is more consistent across a range of tree sizes, and
exhibits a positive relationship with input levels of phylogenetic
signal. However, that relationship is decidedly nonlinear. Thus, neither
\(\lambda\) nor \(\kappa\) are suitable as effect sizes for measuring
the strength of phylogenetic signal, and comparing that signal across
datasets. As an alternative, we propose a standardized effect size based
on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
phylogenetic signal more reliably than does \(\lambda\), and places that
signal on a common scale for statistical comparison. We develop tests
based on \(Z_\kappa\) to provide a mechanism for formally comparing the
strength of phylogenetic signal across datasets, in much the same manner
as effect sizes may be used to summarize patterns in quantitative
meta-analysis. Our approach extends the phylogenetic comparative toolkit
to address hypotheses that compare the strength of phylogenetic signal
between various phenotypic traits, even when those traits are found in
different evolutionary lineages or have different units or scales.
\end{abstract}

@@

By contrast, for Blomberg's \(\kappa\), which is positively ubounded, we
might expect that for any \(\lambda\) used to generate data, estimates
of \(\kappa\) might follow a normal distribution, with values
distributed symmetrically about the input value. This attribute seemed
less reasonable based on the simulations performed by Münkemüller et
al.~(18), which suggested that distributions were positively skewed and
that Blomberg's \(\kappa\) might not behave as a statistic that follows
a normal distribution. However, because their simulations used a
weighted combination of simulated phylogenetic signal strengths, strong
inferences are not possible (and distributional attributes were not the
intended result of their simulations). Thus, for both Pagel's
\(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical moments
across a range of \(\lambda\) used to generate data would be valuable
for adjudicating the relibalilty of these statistics. Furthermore, these
are statistics that appear to have expected values that vary with tree
size (18), making comparisons across studies challenging. Therefore,
transformation of these statistics into \(Z\)-scores would allow
evaluation of the efficacy of each statistic to yield effect sizes that
could be used for comparisons of the strength of phylogenetic signal
across traits and lineages.

@@

\textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
innacurate.} Computer simulations reveal that for \(\hat{\lambda}\), the
distributional expectations of a Bernoulli variable were mostly upheld.
First, the mean value of \(\hat{\lambda}\) increases as \(\lambda\)
increases, but it is negatively-biased (particularly for small tree
sizes), and is consistently less than the input \(\lambda\) value across
most of its range (Fig. 1 black line). Second, the standard deviation of
\(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
smallest at extreme values, implying that the precision in estimating
\(\lambda\) varies across the range of input values (Fig. 1 red line).
Additionally, standard deviations of \(\hat{\lambda}\) are negatively
associated with tree size, and for tress of 128 species or less,
\(\hat{\lambda}\) are quite variable, except for cases when \(\lambda\)
is near or equal to \(1\). Third, the distributions of \(\hat{\lambda}\)
are not normal across its range, but become increasingly skewed at more
extreme values of \(\lambda\) (Fig. 1 blue line). For small tree sizes,
it is also clear that distributions are more platykurtic at intermediate
values of \(\hat{\lambda}\). Taken together these results reveal that
\(\hat\lambda\) inconsistently estimates phylogenetic signal, both
across tree sizes and across the range of input values. Additional
simulations (Supplemental Information) reveal that incorporating
\(\hat\lambda\) in PGLS anova and regression does not adversely affect
the statistical properties of PGLS parameter estimation or model
evaluation (type I error, power, bias in coefficients). Thus, it is
reasonable to incorporate \(\hat\lambda\) in PGLS as a parameter for
tuning the degree of phylogenetic signal in the dependent variables
during the analysis. However, the statistical properties shown in Fig. 1
demonstrate that \(\lambda\) is unsuitable as an effect size for
measuring the strength of phylogenetic signal in data, or for comparing
that signal across datasets.

\begin{figure}
\centering
\includegraphics{new.fig.1.temp.png}
\caption{Response of Pagel's \(\lambda\) to increasing strength of
Brownian motion. Gray line signifies the 1:1 line where the input value
matches the estimate \(\hat\lambda\). At each input level, the dark
black line represents the empirically derived expected value (mean) of
\(\hat\lambda\), the red line is the standard deviation of
\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
skewed distributions).{}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{new.fig.2.temp.png}
\caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of
Brownian motion. Gray line signifies the 1:1 line where the input value
matches the estimate \(\hat\kappa\). At each input level, the dark black
line represents the empirically derived expected value (mean) of
\(\hat\kappa\), the red line is the standard deviation of
\(\hat\kappa\), and the blue line is Shapiro Wilks statistic of
\(\hat\kappa\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
skewed distributions).{}}
\end{figure}

@@

Here we evaluate the ability of \(Z_{\lambda}\) and \(Z_{\kappa}\) to
characterize known levels of phylogenetic signal. Both \(Z_{\lambda}\)
and \(Z_{\kappa}\) are associated with input phylogenetic signal
(\(\lambda\)), indicating that both statistics capture the observed
signal (Fig. 3). However, effect sizes from \(\hat{\lambda}\) made
little sense, as they are more strongly associated with tree size than
they are with the actual phylogenetic signal in the data (Fig. 3). By
contrast, \(Z_{\kappa}\) is much more consistent accross tree sizes, and
increases more linearly with incrasing levels of phylogenetic signal.
Additionally, \(Z_{\kappa}\) exhibits a much stronger association with
phylogenetic signal strength as compared to tree size (Fig. 3), and its
standard deviation across input signal is more consistent. This implies
that similar levels of precision are found with \(Z_{\kappa}\) across
the range of input values. Thus between the two statistics,
\(Z_{\kappa}\) is a more reliable measure of the strength of
phylogenetic signal, and may be used to compare levels of phylogenetic
signal across datasets.

\begin{figure}
\centering
\includegraphics{new.fig.3.temp.png}
\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to
increasing strength of Brownian motion.{}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{new.fig.4.png}
\caption{(A) Linear measures for relative body size, and regions of the
body used to estimate surface area to volume (SA:V) ratios. (B)
Permutation distributions of phylogenetic signal for SA:V and
\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\), with their
95\% confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
\end{figure}

@@

\textbf{Simulations.} Simulations were conducted by generating
pure-birth phylogenies at each of six different tree sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylognetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 100
random trees for each intersection of tree size and \(\lambda\). For
each \(\lambda\) within each tree size, continuous traits were then
simulated on each phylogeny under a BM model of evolution. For each set
of 100 trees we measured the mean values of \(\hat{\lambda}\) and
\(\kappa\), their standard deviation, and calculated the Shapiro-Wilk
\(W\) statistic as a departure from normality (symmetry). For the
latter, a value of \(1.0\) indicates normally distributed values, while
departures from \(1.0\) indicate skewness. Simulations were then
repeated for both bananced and pectinate trees, which yielded
qualitatively similar results (see Supporting Information). Trees
containing polytomies, and an evaluation of \(\hat{\lambda}\) from
models of linear regression and phylogenetic ANOVA, were also
investigated, and results were qualitatively similar to those reported
above (see Supporting Information).

\textbf{Empirical Data.} Surface area to volume ratios (SA:V) and
relative body width (\(\frac{BW}{SVL}\)) measures were obtained from as
species means from individuals of 305 species, from which species means
were obtained (39, 40). A time-dated molecular phylogeny for the group
(45) was pruned to match the species in the phenotypic dataset. The
phylogenetic signal in each trait was then characterized using
\(\kappa\), which was converted to its effect size (\(Z_\kappa\)) using
\texttt{geomorph} 3.3.1 (46, 47), and routines by the authors
(\textbf{to be incorporated in \texttt{geomorph} upon manuscript
acceptance}).

>>>>

\begin{abstract}
Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, however, analytical tools for comparing the
strength of that signal across traits remain largely underdeveloped.
Here we evaluate the efficacy --of-- of Pagel's \(\lambda\) to correctly
estimate the strength of phylogenetic signal in phenotypic traits across
a range of input values. We find that \(\lambda\) behaves as a Bernoulli
random variable, where estimates are increasingly skewed at larger and
smaller input levels of phylogenetic signal. Further, the precision of
\(\lambda\) varies with input signal. Another measure, Blomberg's
\(\kappa\), is more consistent across a range of tree sizes, and
exhibits a positive relationship with input levels of phylogenetic
signal. However, that relationship is decidedly nonlinear. Thus, neither
\(\lambda\) nor \(\kappa\) are suitable as effect sizes for measuring
the strength of phylogenetic signal, and comparing that signal across
datasets. As an alternative, we propose a standardized effect size based
on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
phylogenetic signal more reliably than does \(\lambda\), and places that
signal on a common scale for statistical comparison. We develop tests
based on \(Z_\kappa\) to provide a mechanism for formally comparing the
strength of phylogenetic signal across datasets, in much the same manner
as effect sizes may be used to summarize patterns in quantitative
meta-analysis. Our approach extends the phylogenetic comparative toolkit
to address hypotheses that compare the strength of phylogenetic signal
between various phenotypic traits, even when those traits are found in
different evolutionary lineages or have different units or scales.
\end{abstract}

@@

By contrast, for Blomberg's \(\kappa\), which is positively uNbounded,
we might expect that for any \(\lambda\) used to generate data,
estimates of \(\kappa\) might follow a normal distribution, with values
distributed symmetrically about the input value. This attribute seemed
less reasonable based on the simulations performed by Münkemüller et
al.~(18), which suggested that distributions were positively skewed and
that Blomberg's \(\kappa\) might not behave as a statistic that follows
a normal distribution. However, because their simulations used a
weighted combination of simulated phylogenetic signal strengths, strong
inferences are not possible (and distributional attributes were not the
intended result of their simulations). Thus, for both Pagel's
\(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical moments
across a range of \(\lambda\) used to generate data would be valuable
for adjudicating the --relibalilty-- RELIABILITY of these statistics.
Furthermore, --these are statistics that appear to have expected values
that vary with tree size-- the expected values of these statistics
appear to vary with tree size (18), making comparisons across studies
challenging. Therefore, transformation of these statistics into
\(Z\)-scores would allow evaluation of the efficacy of each statistic to
yield effect sizes that could be used for comparisons of the strength of
phylogenetic signal across traits and lineages.

@@

\textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
innacurate.} Computer simulations reveal that for \(\hat{\lambda}\), the
distributional expectations of a Bernoulli variable were mostly upheld.
First, the mean value of \(\hat{\lambda}\) increases as \(\lambda\)
increases, but it is negatively-biased (particularly for small tree
sizes), and is consistently less than the input \(\lambda\) value across
most of its range (Fig. 1 black line). Second, the standard deviation of
\(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
smallest at extreme values , implying that the precision in estimating
\(\lambda\) varies across the range of input values (Fig. 1 red line).
Additionally, standard deviations of \(\hat{\lambda}\) are negatively
associated with tree size, and for tress of 128 species or less,
\(\hat{\lambda}\) are quite variable, except for cases when \(\lambda\)
is near or equal to \(1\). Third, the distributions of \(\hat{\lambda}\)
are not normal across its range, but become increasingly skewed at more
extreme values of \(\lambda\) (Fig. 1 blue line). For small tree sizes,
it is also clear that distributions are more platykurtic at intermediate
values of \(\hat{\lambda}\). Taken together these results reveal that
\(\hat\lambda\) inconsistently estimates phylogenetic signal, both
across tree sizes and across the range of input values. Additional
simulations (Supplemental Information) reveal that incorporating
\(\hat\lambda\) in PGLS anova and regression does not adversely affect
the statistical properties of PGLS parameter estimation or model
evaluation (type I error, power, bias in coefficients). Thus, it is
reasonable to incorporate \(\hat\lambda\) in PGLS as a parameter for
tuning the degree of phylogenetic signal in the dependent variables
during the analysis. However, the statistical properties shown in Fig. 1
demonstrate that \(\lambda\) is unsuitable as an effect size for
measuring the strength of phylogenetic signal in data, or for comparing
that signal across datasets.

\begin{figure}
\centering
\includegraphics{fig.1.png}
\caption{Response of Pagel's \(\lambda\) to increasing strength of
Brownian motion. Gray line signifies the 1:1 line where the input value
matches the estimate \(\hat\lambda\). At each input level, the dark
black line represents the empirically derived expected value (mean) of
\(\hat\lambda\), the red line is the standard deviation of
\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
skewed distributions).{}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{fig.2.png}
\caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of
Brownian motion. Gray line signifies the 1:1 line where the input value
matches the estimate \(\hat\kappa\). At each input level, the dark black
line represents the empirically derived expected value (mean) of
\(\hat\kappa\), the red line is the standard deviation of
\(\hat\kappa\), and the blue line is Shapiro Wilks statistic of
\(\hat\kappa\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
skewed distributions).{}}
\end{figure}

@@

Here we evaluate the ability of \(Z_{\lambda}\) and \(Z_{\kappa}\) to
characterize known levels of phylogenetic signal. Both \(Z_{\lambda}\)
and \(Z_{\kappa}\) are associated with input phylogenetic signal
(\(\lambda\)), indicating that both statistics capture the observed
signal (Fig. 3). However, effect sizes from \(\hat{\lambda}\) made
little sense, as they are more strongly associated with tree size than
they are with the actual phylogenetic signal in the data (Fig. 3). By
contrast, \(Z_{\kappa}\) is much more consistent accross tree sizes, and
increases more linearly with incrEasing levels of phylogenetic signal.
Additionally, \(Z_{\kappa}\) exhibits a much stronger association with
phylogenetic signal strength as compared to tree size (Fig. 3), and its
standard deviation across input signal is more consistent. This implies
that similar levels of precision are found with \(Z_{\kappa}\) across
the range of input values. Thus between the two statistics,
\(Z_{\kappa}\) is a more reliable measure of the strength of
phylogenetic signal, and may be used to compare levels of phylogenetic
signal across datasets.

\begin{figure}
\centering
\includegraphics{fig.3.png}
\caption{Response of effect sizes \(Z_{\lambda}\) and \(Z_{\kappa}\) to
increasing strength of Brownian motion.{}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{fig.4.png}
\caption{(A) Linear measures for relative body size, and regions of the
body used to estimate surface area to volume (SA:V) ratios. (B)
Permutation distributions of phylogenetic signal for SA:V and
\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\), with their
95\% confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
\end{figure}

@@

\textbf{Simulations.} Simulations were conducted by generating
pure-birth phylogenies at each of six different tree sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylognetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 100
random trees for each intersection of tree size and \(\lambda\). For
each \(\lambda\) within each tree size, continuous traits were then
simulated on each phylogeny under a BM model of evolution. For each set
of 100 trees we measured the mean values of \(\hat{\lambda}\) and
\(\kappa\), their standard deviation, and calculated the Shapiro-Wilk
\(W\) statistic as a departure from normality (symmetry). For the
latter, a value of \(1.0\) indicates normally distributed values, while
departures from \(1.0\) indicate skewness. Simulations were then
repeated for both --bananced-- BALANCED and pectinate trees, which
yielded qualitatively similar results (see Supporting Information).
Trees containing polytomies, and an evaluation of \(\hat{\lambda}\) from
models of linear regression and phylogenetic ANOVA, were also
investigated, and results were qualitatively similar to those reported
above (see Supporting Information).

\textbf{Empirical Data.} Surface area to volume ratios (SA:V) and
relative body width (\(\frac{BW}{SVL}\)) measures were obtained --from
as species means-- from individuals of 305 species, from which species
means were obtained (39, 40). A time-dated molecular phylogeny for the
group (45) was pruned to match the species in the phenotypic dataset.
The phylogenetic signal in each trait was then characterized using
\(\kappa\), which was converted to its effect size (\(Z_\kappa\)) using
\texttt{geomorph} 3.3.1 (46, 47), and routines by the authors
(\textbf{to be incorporated in \texttt{geomorph} upon manuscript
acceptance}).

### 2020-08-12 Manuscript/PNAS-Adams-02.Rmd
vsn 2

abstract: |
  Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, however, analytical tools for comparing the strength of that signal across traits remain largely underdeveloped. Here we evaluate the efficacy --of-- of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values. We find that $\lambda$ behaves as a Bernoulli random variable, where estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal. Further, the precision of $\lambda$ varies with input signal. Another measure, Blomberg's $\kappa$, is more consistent across a range of tree sizes, and exhibits a positive relationship with input levels of phylogenetic signal. However, that relationship is decidedly nonlinear. Thus, neither $\lambda$ nor $\kappa$ are suitable as effect sizes for measuring the strength of phylogenetic signal, and comparing that signal across datasets. As an alternative, we propose a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison. We develop tests based on $Z_\kappa$ to provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis. Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.

@@

![Response of Pagel's $\lambda$ to increasing strength of Brownian motion. Gray line signifies the 1:1 line where the input value matches the estimate $\hat\lambda$. At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\lambda$, the red line is the standard deviation of $\hat\lambda$, and the blue line is Shapiro Wilks statistic of $\hat\lambda$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f1"></span>](new.fig.1.temp.png)

@@

![Response of Blomberg's $\hat\kappa$ to increasing strength of Brownian motion. Gray line signifies the 1:1 line where the input value matches the estimate $\hat\kappa$. At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\kappa$, the red line is the standard deviation of $\hat\kappa$, and the blue line is Shapiro Wilks statistic of $\hat\kappa$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f2"></span>](new.fig.2.temp.png)

@@

![Response of effect sizes $Z_{\lambda}$ and $Z_{\kappa}$ to increasing strength of Brownian motion.<span data-label="fig:f3"></span>](new.fig.3.temp.png)

@@

![(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios. (B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars. (C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](new.fig.4.png)

>>>>

abstract: |
  Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, however, analytical tools for comparing the strength of that signal across traits remain largely underdeveloped. Here we evaluate the efficacy of Pagel's $\lambda$ to correctly estimate the strength of phylogenetic signal in phenotypic traits across a range of input values. We find that $\lambda$ behaves as a Bernoulli random variable, where estimates are increasingly skewed at larger and smaller input levels of phylogenetic signal. Further, the precision of $\lambda$ varies with input signal. Another measure, Blomberg's $\kappa$, is more consistent across a range of tree sizes, and exhibits a positive relationship with input levels of phylogenetic signal. However, that relationship is decidedly nonlinear. Thus, neither $\lambda$ nor $\kappa$ are suitable as effect sizes for measuring the strength of phylogenetic signal, and comparing that signal across datasets. As an alternative, we propose a standardized effect size based on $\kappa$, ($Z_\kappa$), which measures the strength of phylogenetic signal more reliably than does $\lambda$, and places that signal on a common scale for statistical comparison. We develop tests based on $Z_\kappa$ to provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis. Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal between various phenotypic traits, even when those traits are found in different evolutionary lineages or have different units or scales.

@@

![Response of Pagel's $\lambda$ to increasing strength of Brownian motion. Gray line signifies the 1:1 line where the input value matches the estimate $\hat\lambda$. At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\lambda$, the red line is the standard deviation of $\hat\lambda$, and the blue line is Shapiro Wilks statistic of $\hat\lambda$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f1"></span>](fig.1.png)

@@

![Response of Blomberg's $\hat\kappa$ to increasing strength of Brownian motion. Gray line signifies the 1:1 line where the input value matches the estimate $\hat\kappa$. At each input level, the dark black line represents the empirically derived expected value (mean) of $\hat\kappa$, the red line is the standard deviation of $\hat\kappa$, and the blue line is Shapiro Wilks statistic of $\hat\kappa$ ($W=1.0$ signifies normality, $W< 1.0$ represent skewed distributions).<span data-label="fig:f2"></span>](fig.2.png)

@@

![Response of effect sizes $Z_{\lambda}$ and $Z_{\kappa}$ to increasing strength of Brownian motion.<span data-label="fig:f3"></span>](fig.3.png)

@@

![(A) Linear measures for relative body size, and regions of the body used to estimate surface area to volume (SA:V) ratios. (B) Permutation distributions of phylogenetic signal for SA:V and $\frac{BW}{SVL}$, with observed values shown as vertical bars. (C) Effect sizes ($Z_\kappa$) for SA:V and $\frac{BW}{SVL}$, with their 95% confidence intervals (CI not standardized by $\sqrt(n)$).<span data-label="fig:f4"></span>](fig.4.png)

### 2020-08-12 Manuscript/PNAS-Adams-02.tex
vsn 2

\begin{abstract}
Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, however, analytical tools for comparing the
strength of that signal across traits remain largely underdeveloped.
Here we evaluate the efficacy --of-- of Pagel's \(\lambda\) to correctly
estimate the strength of phylogenetic signal in phenotypic traits across
a range of input values. We find that \(\lambda\) behaves as a Bernoulli
random variable, where estimates are increasingly skewed at larger and
smaller input levels of phylogenetic signal. Further, the precision of
\(\lambda\) varies with input signal. Another measure, Blomberg's
\(\kappa\), is more consistent across a range of tree sizes, and
exhibits a positive relationship with input levels of phylogenetic
signal. However, that relationship is decidedly nonlinear. Thus, neither
\(\lambda\) nor \(\kappa\) are suitable as effect sizes for measuring
the strength of phylogenetic signal, and comparing that signal across
datasets. As an alternative, we propose a standardized effect size based
on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
phylogenetic signal more reliably than does \(\lambda\), and places that
signal on a common scale for statistical comparison. We develop tests
based on \(Z_\kappa\) to provide a mechanism for formally comparing the
strength of phylogenetic signal across datasets, in much the same manner
as effect sizes may be used to summarize patterns in quantitative
meta-analysis. Our approach extends the phylogenetic comparative toolkit
to address hypotheses that compare the strength of phylogenetic signal
between various phenotypic traits, even when those traits are found in
different evolutionary lineages or have different units or scales.
\end{abstract}

>>>>

\begin{abstract}
Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, however, analytical tools for comparing the
strength of that signal across traits remain largely underdeveloped.
Here we evaluate the efficacy of Pagel's \(\lambda\) to correctly
estimate the strength of phylogenetic signal in phenotypic traits across
a range of input values. We find that \(\lambda\) behaves as a Bernoulli
random variable, where estimates are increasingly skewed at larger and
smaller input levels of phylogenetic signal. Further, the precision of
\(\lambda\) varies with input signal. Another measure, Blomberg's
\(\kappa\), is more consistent across a range of tree sizes, and
exhibits a positive relationship with input levels of phylogenetic
signal. However, that relationship is decidedly nonlinear. Thus, neither
\(\lambda\) nor \(\kappa\) are suitable as effect sizes for measuring
the strength of phylogenetic signal, and comparing that signal across
datasets. As an alternative, we propose a standardized effect size based
on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
phylogenetic signal more reliably than does \(\lambda\), and places that
signal on a common scale for statistical comparison. We develop tests
based on \(Z_\kappa\) to provide a mechanism for formally comparing the
strength of phylogenetic signal across datasets, in much the same manner
as effect sizes may be used to summarize patterns in quantitative
meta-analysis. Our approach extends the phylogenetic comparative toolkit
to address hypotheses that compare the strength of phylogenetic signal
between various phenotypic traits, even when those traits are found in
different evolutionary lineages or have different units or scales.
\end{abstract}
